Towards Proof-Reconstruction of Problems in Classical Propositional Logic

### 2017-10-15 paper/sections/proof-reconstruction.tex
[ paper ] first description version for canonicalize.

\subsubsection{A practical Example.}
Let us consider the following example to exhibit the technique
mentioned above for an inner function present in the emulated version
of the \strip inference.

@@

We emulate each \Metis inference rule by proving a
function with the same name, and a theorem that
proves the soundness of the function.
The pattern is the following.

\begin{itemize}
\item \texttt{rule} is the \Metis inference rule.
\item $rule$ is the emulated version function for \texttt{rule} inference.
\begin{align*}
\begin{split}
&rule : \Prop → \Prop → \Prop\\
&rule(φ, ψ) =
  \begin{cases}
  ψ, &\text{if }rule\text{ built }ψ\text{ by applying valid transformations to }φ\\
  φ, &\text{otherwise.}
  \end{cases}
\end{split}
\end{align*}

@@

To prove the correctness of the process mentioned above, we state
the Theorem~\ref{thm:thm-strip}. But first, let us prove some
lemmas for $unshunt$ function presented in Eq.~(\ref{eq:unshunt-fixed}),
and $split$ function from the definition in Eq.~(\ref{eq:split-definition}).

@@

\begin{proof} Use induction on the cases defined by the outcome of the
$unshunt$ function in Eq.~(\ref{eq:unshunt-structured}).
\begin{itemize}
	\item When $n = 0$, by definition, $Γ ⊢ φ$.
	\item When $n = 1$, we apply the following theorem from~\cite{AgdaProp},
		\begin{equation*}
		\texttt{∧⇒-to-⇒⇒}\ :\  Γ ⊢ (φ₁ ∧ φ₂) ⇒ φ₃ → Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃).
		\end{equation*}
  \item For $n > 1$, we will use induction on the structure of the second
			  argument.

\begin{itemize}
\item $(φ₁ ⇒ (φ₂ ⇒ φ₃))$
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ unshunt(φ₁ ⇒ (φ₂ ⇒ φ₃), n)$}
  \RightLabel{axiom}
  \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ (φ₂ ⇒ φ₃), n-1)$}
  \RightLabel{lem-unshunt}
  \UnaryInfC{Γ ⊢ (φ₁ ∧ φ₂) ⇒ φ₃}
  \RightLabel{∧⇒-to-⇒⇒.}
  \UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃)}
  \end{bprooftree}
\end{equation*}

@@

Finally, using the theorem \texttt{⇒∧⇒-to-⇒∧} from~\cite{AgdaProp},
\begin{equation*}
  \texttt{⇒∧⇒-to-⇒∧}\ :\ Γ ⊢ (φ₁ ⇒ φ₂) ∧ (φ₁ ⇒ φ₃) → Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃),
\end{equation*}
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}_1$}
  \AxiomC{$\mathcal{D}_2$}
  \RightLabel{∧-intro}
  \BinaryInfC{$Γ ⊢ (φ₁ ⇒ φ₂) ∧ (φ₁ ⇒ φ₃)$}
  \RightLabel{\tt ⇒∧⇒-to-⇒∧.}
  \UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃)}
  \end{bprooftree}
\end{equation*}

\end{itemize}
\end{itemize}
\end{proof}

@@

The definition above suffers termination problems since it is not
a structural recursion. Therefore, we have reformulated the equation above to solve such a issue following the technique
described in Section~\ref{ssec:structural-recursion}.
Here and below, we refer to $split$ as the function after applying
the process mentioned above.

@@

\begin{proof} We now proceed by induction on the structure of the
formula $φ$ in the definition of Eq.~(\ref{eq:split-definition}).
For a complete proof, we refer the reader to~\cite{AgdaMetis}.\\[3mm]

\begin{itemize}
\item $(φ ≡ φ₁ ∧ φ₂)$. We get a proof for each conjunct and using the introduction rule for conjunction connective, the result follows.

@@

\begin{equation*}
(\mathcal{D})\hspace{3mm}
\begin{bprooftree}
\AxiomC{$Γ ⊢ \mysplit(φ, n)$}
\RightLabel{axiom}
\UnaryInfC{$Γ ⊢ unshunt(¬ φ₁) ∧ unshunt(¬ φ₁ ⇒ \mysplit(φ₂, n))$}
\RightLabel{∧-proj₁}
\UnaryInfC{$Γ ⊢ unshunt(¬ φ₁)$}
\RightLabel{lem-unshunt,}
\UnaryInfC{$Γ ⊢ ¬ φ₁$}
\end{bprooftree}
\end{equation*}
\end{itemize}
\end{proof}

@@

First, we define the $rdisj$ function in Eq.~(\ref{eq:rdisj-def}) to convert a disjunction in its right-associative form.
\begin{definition}[rdisj]
\begin{align}
  \begin{split}
    \label{eq:rdisj-def}
    &rdisj(φ) : \Prop \to \Prop\\
    &rdisj(φ) =
    \begin{cases}
    rdisj(φ₁ ∨ (φ₂ ∨ φ₃)),  & \text{if } φ ≡ (φ₁ ∨ φ₂) ∨ φ₃\\
    φ₁ ∨ rdisj(φ₂),         & \text{if } φ ≡ φ₁ ∨ φ₂\\
    φ,                      & \text{otherwise.}
    \end{cases}
  \end{split}
\end{align}
\end{definition}

@@

\begin{lemma}[lem-sbuild$_{∨}$]
\label{lem:lem-sbuild-or}
$Γ ⊢ φ$, and $ψ \in\Prop$ then $Γ ⊢ sbuild_{∨}(φ, ψ)$.
\end{lemma}

@@

\begin{theorem}[thm-reorder$_{∨}$]
  \label{thm-reorder-or}
  $Γ ⊢ φ$, and $ψ\in\Prop$ then $Γ ⊢ reorder_{∨}(φ, ψ)$.
\end{theorem}

@@

\begin{lemma}[lem-reorder$_{∧}$]
  \label{lem:lem-reorder-and}
  $Γ ⊢ φ$, and $ψ \in\Prop$ then $Γ ⊢ reorder(φ, ψ)$.
\end{lemma}

@@

\begin{lemma}[lem-conjunct$_{∨}$]
  \label{lem:lem-conjunct-or}
  $Γ ⊢ φ$, and $ψ \in\Prop$ then $Γ ⊢ conjunct_{∨}(φ, ψ)$.
\end{lemma}

@@

\begin{theorem}[thm-reorder$_{∧∨}$]
  \label{thm:thm-reorder-and-or}
  $Γ ⊢ φ$, and $ψ \in\Prop$ then $Γ ⊢ reorder_{∧∨}(φ, ψ)$.
\end{theorem}

@@

\begin{theorem}[thm-resolve]
  \label{thm:thm-resolve}
  $Γ ⊢ φ₁$, $Γ ⊢ φ₂$, a literal $l$, and $ψ\in\Prop$ then $Γ ⊢ resolve(φ₁, φ₂, l, ψ)$.
\end{theorem}

\begin{proof}\hspace{3mm}
\begin{itemize}
 \item
 \begin{equation*}
    \begin{bprooftree}
      \AxiomC{$Γ ⊢ φ₁$}
      \RightLabel{thm-reorder$_{∨}$.}
      \UnaryInfC{$Γ ⊢ reorder_{∨}(φ₁, l ∨ ψ)$}
   \end{bprooftree}
  \end{equation*}
\item
 \begin{equation*}
    \begin{bprooftree}
      \AxiomC{$Γ ⊢ φ₂$}
      \RightLabel{thm-reorder$_{∨}$.}
      \UnaryInfC{$Γ ⊢ reorder_{∨}(φ₂, ¬ l ∨ ψ)$}
   \end{bprooftree}
  \end{equation*}
  \item
  \begin{equation*}
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ reorder_{∨}(φ₁, l ∨ ψ)$}
    \AxiomC{$Γ ⊢ reorder_{∨}(φ₂, ¬ l ∨ ψ)$}
    \RightLabel{∧-intro}
    \BinaryInfC{$Γ ⊢ reorder_{∨}(φ₁, l ∨ ψ) ∧ reorder_{∨}(φ₂, ¬ l ∨ ψ)$}
    \RightLabel{lem-rsol.}
    \UnaryInfC{$Γ ⊢ rsol(reorder_{∨}(φ₁, l ∨ ψ) ∧ reorder_{∨}(φ₂, ¬ l ∨ ψ))$}
  \end{bprooftree}
  \end{equation*}
\end{itemize}
\end{proof}

@@

\subsubsection{Normalization.}
\label{sssec:normal_forms}
\change{Here lacks of an introduction!}
A negative normal form for the formula φ is one in which negations
appear only in the literals and the expression is in terms only of
the boolean operators (¬, ∧, ∨).

% To transform a formula φ to its equivalent \abbre{NNF} form, we
% apply recursively the following equations. The function type for
% \name{nnf} is $\Prop \to \Prop$.

% \begin{lemma}[lem-nnf]
%   \label{lem:lem-nnf}
%   $Γ ⊢ φ$ then $Γ ⊢ nnf(φ)$, where,
%   \begin{align*}
%     \begin{split}
%     \label{eq:nnf-definition}
%     nnf(φ) =
%       \begin{cases}
%       nnf(φ₁) ∧ nnf(φ₂),
%          &\text{if }φ ≡ φ₁ ∧ φ₂ \\
%       nnf(φ₁) ∨ nnf(φ₂),
%          &\text{if }φ ≡ φ₁ ∨ φ₂ \\
%       nnf(¬ φ₁ ∨ φ₂),
%          &\text{if }φ ≡ φ₁ ⇒ φ₂ \\
%       nnf(φ₁ ⇒ φ₂ ∧ φ₂ ⇒ φ₁),
%          &\text{if }φ ≡ φ₁ ⇔ φ₂ \\
%       nnf(¬ φ₁ ∨ ¬ φ₂),
%          &\text{if }φ ≡ ¬ (φ₁ ∧ φ₂) \\
%       nnf(¬ φ₁ ∧ ¬ φ₂),
%          &\text{if }φ ≡ ¬ (φ₁ ∨ φ₂) \\
%       nnf(φ₁),
%          &\text{if }φ ≡ ¬ ¬ φ₁ \\
%       nnf(¬ (φ₂ ∨ ¬ φ₁)),
%          &\text{if }φ ≡ ¬ (φ₁ ⇒ φ₂) \\
%       nnf(¬ (φ₁ ⇒ φ₂ ∧ φ₂ ⇒ φ₁)),
%          &\text{if }φ ≡ ¬ (φ₁ ⇔ φ₂) \\
%       ⊥,  &\text{if }φ ≡ ¬ ⊤ \\
%       ⊤,  &\text{if }φ ≡ ¬ ⊥ \\
%       φ,  & \text{otherwise}.
%       \end{cases}
%     \end{split}
%   \end{align*}
% \end{lemma}

% \begin{lemma}[lem-dist$_{∨}$]
%   \label{lem:lem-dist-or}
%   $Γ ⊢ φ$ then $Γ ⊢ dist_{∨}(φ)$, where,
%   \begin{align*}
%       \begin{split}
%       dist_{∨}(φ) =
%         \begin{cases}
%           dist_{∨}(φ₁ ∨ φ₂) ∧ dist_{∨}(φ₂ ∨ φ₃),
%             &\text{if }φ ≡ (φ₁ ∧ φ₂) ∨ φ₃\\
%           dist_{∨}(φ₁ ∨ φ₃) ∧ dist_{∨}(φ₁ ∨ φ₃),
%             &\text{if }φ ≡ φ₁ ∨ (φ₂ ∧ φ₃)\\
%           φ, &\text{otherwise.}
%         \end{cases}
%       \end{split}
%   \end{align*}
% \end{lemma}

% \begin{lemma}[lem-dist]
%   \label{lem:lem-dist}
%   $Γ ⊢ φ$ then $Γ ⊢ dist(φ)$, where,
%   \begin{align*}
%       \begin{split}
%       dist(φ) =
%         \begin{cases}
%           dist(φ₁) ∧ dist(φ₂),
%             &\text{if }φ ≡ φ₁ ∧ φ₂\\
%           dist_{∨}(dist(φ₁) ∨ dist(φ₂)),
%             &\text{if }φ ≡ φ₁ ∨ φ₂\\
%           φ, &\text{otherwise.}
%         \end{cases}
%       \end{split}
%   \end{align*}
% \end{lemma}

% \begin{sketchproof}
% Case when $φ ≡ φ₁ ∨ φ₂$.\\
% For $i = 1, 2$,
% \begin{equation}
%   \begin{bprooftree}
%     \RightLabel{assume}
%     \AxiomC{$Γ, φ_{i} ⊢ φ_{i}$}
%     \RightLabel{lem-dist}
%     \UnaryInfC{$Γ ⊢ dist(φ_{i})$}
%     \RightLabel{∨-intro$_{1,2}$}
%     \UnaryInfC{$Γ, φ_{i} ⊢ dist(φ₁) ∨ dist(φ₂)$}
%     \RightLabel{lem-dist$_{∨}$}
%     \UnaryInfC{$Γ, φ_{i} ⊢ dist_{∨}(dist(φ₁) ∨ dist(φ₂))$.}
%     \end{bprooftree}
% \end{equation}
% Then, by using the ∨-elim rule, from the last derivation above we derive
% $Γ, φ₁ ∨ φ₂ ⊢ dist_{∨}(dist(φ₁) ∨ dist(φ₂))$, and the lemma follows.
% \end{sketchproof}

% \begin{theorem}[thm-cnf]
% \label{thm:thm-cnf}
%   $Γ ⊢ φ$ then $Γ ⊢ cnf(φ)$, where,
%   \begin{equation}
%     cnf(φ) = (dist \circ nnf) (\varphi).
%   \end{equation}
% \end{theorem}

% \begin{proof}
%   Composition of the Lemma \ref{lem:lem-dist} and Lemma \ref{lem:lem-nnf}.
% \end{proof}
% % subsubsection normal_forms (end)
% % --------------------------------------------------------------------

>>>>

\subsubsection{A Practical Example.}
Let us consider the following example to exhibit the technique
mentioned above for an inner function present in the emulated version
of the \strip inference.

@@

In this section, we emulate each \Metis inference rule by proving a
function with the same name, and a theorem that
proves the soundness of the function. Then, we will find
the following pattern to present the formalization for
the \Metis inference rules.

\begin{itemize}
\item \texttt{rule} is a \Metis inference rule
\item $rule$ is the emulated version function for
the \texttt{rule} inference.
\begin{align*}
\begin{split}
&rule : \Prop → \Prop → \Prop\\
&rule(φ, ψ) =
  \begin{cases}
  ψ, &\text{if }rule\text{ built }ψ\text{ by applying valid transformations to }φ\\
  φ, &\text{otherwise.}
  \end{cases}
\end{split}
\end{align*}

@@

To prove the correctness of the process mentioned above, we state
the Theorem~\ref{thm:thm-strip}. But first, let us prove some
lemmas for $unshunt$ function presented in Def.~(\ref{def:unshunt}),
and for $split$ function presented in Eq.~(\ref{eq:split-definition}).

@@

\begin{sketchproof} Use induction on the cases defined by the outcome of the
$unshunt$ function in Eq.~(\ref{eq:unshunt-structured}).
\begin{itemize}
	\item If $n = 0$, by definition we conclude $Γ ⊢ φ$.
	\item If $n = 1$, we apply the following theorem proved in~\cite{AgdaProp},
		\begin{equation*}
		\texttt{∧⇒-to-⇒⇒}\ :\  Γ ⊢ (φ₁ ∧ φ₂) ⇒ φ₃ → Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃).
		\end{equation*}
  \item For $n > 1$, we will use induction on the structure of the second
			  argument.
\begin{itemize}
\item $(φ₁ ⇒ (φ₂ ⇒ φ₃))$
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ unshunt(φ₁ ⇒ (φ₂ ⇒ φ₃), n)$}
  \RightLabel{axiom}
  \UnaryInfC{$Γ ⊢ unshunt(φ₁ ⇒ (φ₂ ⇒ φ₃), n-1)$}
  \RightLabel{lem-unshunt}
  \UnaryInfC{Γ ⊢ (φ₁ ∧ φ₂) ⇒ φ₃}
  \RightLabel{∧⇒-to-⇒⇒.}
  \UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ⇒ φ₃)}
  \end{bprooftree}
\end{equation*}

@@

Finally, using the theorem \texttt{⇒∧⇒-to-⇒∧} from~\cite{AgdaProp},
\begin{equation*}
  \texttt{⇒∧⇒-to-⇒∧}\ :\ Γ ⊢ (φ₁ ⇒ φ₂) ∧ (φ₁ ⇒ φ₃) → Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃),
\end{equation*}
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}_1$}
  \AxiomC{$\mathcal{D}_2$}
  \RightLabel{∧-intro}
  \BinaryInfC{$Γ ⊢ (φ₁ ⇒ φ₂) ∧ (φ₁ ⇒ φ₃)$}
  \RightLabel{\tt ⇒∧⇒-to-⇒∧.}
  \UnaryInfC{Γ ⊢ φ₁ ⇒ (φ₂ ∧ φ₃)}
  \end{bprooftree}
\end{equation*}
\item Other cases are proved in a similar way.
\end{itemize}
\end{itemize}
\end{sketchproof}

@@

The definition above suffers termination problems since it is not
a structural recursion. Therefore, we have reformulated the equation
above to solve such a issue following the technique
described in Section~\ref{ssec:structural-recursion}. We refer to $split$
as the function after applying the process mentioned above here and below
unless otherwise stated.

@@

\begin{sketchproof}
We now proceed by induction on the structure of the
formula $φ$ by following the cases defined in
Eq.~(\ref{eq:split-definition}).
We present some cases but for a complete proof, we refer the reader to~\cite{AgdaMetis}.
\begin{itemize}
\item $(φ ≡ ¬ ¬ φ₁)$
%Using the double negation equivalence, we conclude.
\begin{equation*}
  \begin{bprooftree}
 \AxiomC{$Γ ⊢ \mysplit(φ,n)$}
  \RightLabel{axiom}
  \UnaryInfC{$Γ ⊢ unshunt(split(¬¬φ₁,n-1))$}
  \RightLabel{lem-unshunt}
  \UnaryInfC{$Γ ⊢split(¬¬φ₁,n-1)$}
  \RightLabel{\thmsplit}
  \UnaryInfC{$Γ ⊢ ¬¬φ₁$}
  \RightLabel{axiom.}
  \UnaryInfC{$Γ ⊢ φ$}
  \end{bprooftree}
\end{equation*}

\item $(φ ≡ φ₁ ∧ φ₂)$. We get a proof for each conjunct and using the introduction rule for conjunction connective, the result follows.

@@

\begin{equation*}
(\mathcal{D})\hspace{3mm}
\begin{bprooftree}
\AxiomC{$Γ ⊢ \mysplit(φ, n)$}
\RightLabel{axiom}
\UnaryInfC{$Γ ⊢ unshunt(¬ φ₁) ∧ unshunt(¬ φ₁ ⇒ \mysplit(φ₂, n))$}
\RightLabel{∧-proj₁}
\UnaryInfC{$Γ ⊢ unshunt(¬ φ₁)$}
\RightLabel{lem-unshunt,}
\UnaryInfC{$Γ ⊢ ¬ φ₁$}
\end{bprooftree}
\end{equation*}
\end{itemize}
\end{sketchproof}

@@

First, we define the $rdisj$ function in Eq.~(\ref{eq:rdisj-def}) to convert a disjunction in its right-associative form.

\begin{definition}[rdisj]
\label{def:rdisj}
\begin{align}
  \begin{split}
  \label{eq:rdisj-def}
    &rdisj(φ) : \Prop \to \Prop\\
    &rdisj(φ) =
    \begin{cases}
    rdisj(φ₁ ∨ (φ₂ ∨ φ₃)),  & \text{if } φ ≡ (φ₁ ∨ φ₂) ∨ φ₃\\
    φ₁ ∨ rdisj(φ₂),         & \text{if } φ ≡ φ₁ ∨ φ₂\\
    φ,                      & \text{otherwise.}
    \end{cases}
  \end{split}
\end{align}
\end{definition}

@@

\begin{lemma}[lem-sbuild$_{∨}$]
\label{lem:lem-sbuild-or}
$Γ ⊢ φ$, and $ψ : \Prop$ then $Γ ⊢ sbuild_{∨}(φ, ψ)$.
\end{lemma}

@@

\begin{theorem}[thm-reorder$_{∨}$]
  \label{thm-reorder-or}
  $Γ ⊢ φ$, and $ψ: \Prop$ then $Γ ⊢ reorder_{∨}(φ, ψ)$.
\end{theorem}

@@

\begin{lemma}[lem-reorder$_{∧}$]
  \label{lem:lem-reorder-and}
  $Γ ⊢ φ$, and $ψ : \Prop$ then $Γ ⊢ reorder(φ, ψ)$.
\end{lemma}

@@

\begin{lemma}[lem-conjunct$_{∨}$]
  \label{lem:lem-conjunct-or}
  $Γ ⊢ φ$, and $ψ : \Prop$ then $Γ ⊢ conjunct_{∨}(φ, ψ)$.
\end{lemma}

@@

\begin{theorem}[thm-reorder$_{∧∨}$]
  \label{thm:thm-reorder-and-or}
  $Γ ⊢ φ$, and $ψ : \Prop$ then $Γ ⊢ reorder_{∧∨}(φ, ψ)$.
\end{theorem}

@@

\begin{theorem}[thm-resolve]
  \label{thm:thm-resolve}
  $Γ ⊢ φ₁$, $Γ ⊢ φ₂$, a literal $l$, and $ψ: \Prop$ then $Γ ⊢ resolve(φ₁, φ₂, l, ψ)$.
\end{theorem}
\Needspace{3\baselineskip}
\begin{proof}
 \begin{equation*}
    \begin{bprooftree}
      \AxiomC{$Γ ⊢ φ₁$}
      \RightLabel{thm-reorder$_{∨}$,}
      \UnaryInfC{$Γ ⊢ reorder_{∨}(φ₁, l ∨ ψ)$}
   \end{bprooftree}
  \end{equation*}
    \vskip 1.5mm
 \begin{equation*}
    \begin{bprooftree}
      \AxiomC{$Γ ⊢ φ₂$}
      \RightLabel{thm-reorder$_{∨}$,}
      \UnaryInfC{$Γ ⊢ reorder_{∨}(φ₂, ¬ l ∨ ψ)$}
   \end{bprooftree}
  \end{equation*}
  \vskip 1.5mm
  \begin{equation*}
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ reorder_{∨}(φ₁, l ∨ ψ)$}
    \AxiomC{$Γ ⊢ reorder_{∨}(φ₂, ¬ l ∨ ψ)$}
    \RightLabel{∧-intro}
    \BinaryInfC{$Γ ⊢ reorder_{∨}(φ₁, l ∨ ψ) ∧ reorder_{∨}(φ₂, ¬ l ∨ ψ)$}
    \RightLabel{lem-rsol.}
    \UnaryInfC{$Γ ⊢ rsol(reorder_{∨}(φ₁, l ∨ ψ) ∧ reorder_{∨}(φ₂, ¬ l ∨ ψ))$}
  \end{bprooftree}
  \end{equation*}
\end{proof}

@@

\subsubsection{Normalization.}
\label{sssec:normal_forms}
In the core of \Metis, we find an important module:
\texttt{Normalize.sml}.
This module is the responsible for rules like \canonicalize, \clausify,
\conjunct or \simplify.
One of the main purposes of the module consists of adding the axioms to the problem. Into this process, some transformation are carried out by different inference rules.
For example, the \canonicalize rule can convert a FOF formula to their CNF form. In that process, the formula can suffers different transformations by removing redundancies (tautologies or definitions), converting the proposition to its normalized negative normal form, to its conjunctive normal form, or to a specialized normal form.

\begin{definition}[\abbre{NProp}]
  \NProp is the type for a formula in normalized negative normal form
  in which negations appear only in the literals and the expression is in terms only of ($⊥$, $⊤$, $¬$, $∧$, $∨$, $⊻$) connectives. The $⊻$ connective is the exclusive disjunction.
  We define the \NProp type in a similar way as we did for \Prop type.
\end{definition}

A sequent $Γ ⟝ φ$ represents a theorem where
$Γ$ is a set of \Prop propositions premises, and $φ : \NProp$ is the
sequent's conclusion. These theorems have a subset of rules from
the classical inference rules in Fig.~\ref{fig:CPL-inference-rules}
that apply for the connectives that coincides with \NProp type.

% φ ⊻ ψ if and only if ¬ (φ ⇔ ψ).

\begin{figure}
  \[
    \begin{bprooftree}
      \AxiomC{$φ ∧ ⊥$}
      \UnaryInfC{$⊥$}
    \end{bprooftree}
    \qquad
    \begin{bprooftree}
      \AxiomC{$φ ∧ ⊤$}
      \UnaryInfC{$φ$}
    \end{bprooftree}
    \qquad
    \begin{bprooftree}
      \AxiomC{$φ ∧ ¬ φ$}
      \UnaryInfC{$⊥$}
    \end{bprooftree}
    \qquad
    \begin{bprooftree}
      \AxiomC{$φ ∧ φ$}
      \UnaryInfC{$φ$}
    \end{bprooftree}
  \]

  \[
    \begin{bprooftree}
      \AxiomC{$φ ∨ ⊥$}
      \UnaryInfC{$φ$}
    \end{bprooftree}
    \qquad
    \begin{bprooftree}
      \AxiomC{$φ ∨ ⊤$}
      \UnaryInfC{$⊤$}
    \end{bprooftree}
    \qquad
    \begin{bprooftree}
      \AxiomC{$φ ∨ ¬ φ$}
      \UnaryInfC{$⊤$}
    \end{bprooftree}
    \qquad
    \begin{bprooftree}
      \AxiomC{$φ ∨ φ$}
      \UnaryInfC{$φ$}
    \end{bprooftree}
  \]
\caption{Some definitions applied to translate a formula into its
normalized normal form.}
\label{fig:conjunctive-disjunctive-simpl}
\end{figure}

% -------------------------------------------------------------------
\subsubsection{Redundancy.}
\label{sssec:redundancy}

In this section we provide the functions and theorems to
remove inside a normalized formula the definitions showed in
Fig.~\ref{fig:conjunctive-disjunctive-simpl}.

\begin{remark}
We assume all formulas are in right-associative form in
the input of the functions by applying a version of $rdisj$ or $rconj$
functions for normalized propositions (see Def.~\ref{def:rdisj}).
\end{remark}

In a disjunction, $φ = φ₁ ∨ φ₂ ∨ \cdots ∨ φₙ$, we say $ψ ∈_{∨} φ$,
if there is some $i = 1, \cdots, n$ such that $ψ = φᵢ$.
Note that $ψ ∈_{∨} φ$ can be defined as the equality
$ψ ≡ conjunct_{∨}(φ, ψ)$.

\begin{definition}[rm$_∨$]
  \label{def:rm-or}
  \begin{align*}
    \begin{split}
      &rm_{∨} :  \NProp \to \NProp\\
      &rm_{∨}(φ) =
      \begin{cases}
        rm_{∨}(φ₂), &\text{if }φ ≡ φ₁ ∨ φ₂,\ φ₁ ∈_{∨} φ₂\\
        φ₁ ∨ rm_{∨}(φ₂), &\text{if }φ ≡ φ₁ ∨ φ₂\\
        φ,  &\text{otherwise.}
      \end{cases}
    \end{split}
  \end{align*}
\end{definition}

\begin{lemma}[lem-rm$_{∨}$]
  \label{lem:lem_rm-or}
  $Γ ⟝ φ$ then $Γ ⟝ rm_{∨}(φ)$.
\end{lemma}

Now, we have removed redundancies in the disjunctions by applying the
$rm_{∨}$, we can define a similar function to work with conjunctions.

\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$Γ ⟝ φ ∨ φ$}
  \RightLabel{lem-rm$_∨$,}
  \UnaryInfC{$Γ ⟝ φ$}
  \end{bprooftree}
  \qquad
  \begin{bprooftree}
  \AxiomC{$Γ ⟝ φ ∧ φ$}
  \RightLabel{lem-rm$_∧$.}
  \UnaryInfC{$Γ ⟝ φ$}
  \end{bprooftree}
\end{equation*}

In a conjunction, $φ = φ₁ ∧ φ₂ ∧ \cdots ∧ φₙ$, we say
$ψ ∈_{∧} φ$, if there is some $i = 1, \cdots, n$ such that $ψ = φᵢ$.
We define $ψ ∈_{∧} φ$ as the equality $ψ ≡ conjunct(φ, ψ)$.

\begin{definition}[rm$_∧$]
\label{eq:rm-and}
  \begin{align*}
    \begin{split}
    &rm_{∧} : \NProp \to \NProp\\
    &rm_{∧}(φ) =
    \begin{cases}
      rm_{∧}(φ₂),      &\text{if }φ ≡ φ₁ ∧ φ₂, φ₁ ∈_{∧} φ₂\\
      φ₁ ∧ rm_{∧}(φ₂), &\text{if }φ ≡ φ₁ ∧ φ₂\\
      φ,               &\text{otherwise.}
    \end{cases}
    \end{split}
  \end{align*}
\end{definition}

\begin{lemma}[lem-rm$_{∧}$] % (fold)
  \label{lem:lem_rm-and}
  $Γ ⟝ φ$ then $Γ ⟝ rm_{∧}(φ)$.
\end{lemma}

Continuing with the rest of tautologies from
Fig.~\ref{fig:conjunctive-disjunctive-simpl}, we define
$ndisj_{∨}$ function and $ndisj_{∨}$ function that
take into account the following cases.

\begin{equation*}
\begin{bprooftree}
  \AxiomC{$Γ ⟝ φ ∨ ¬~φ$}
  \RightLabel{lem-ndisj$_{∨}$,}
  \UnaryInfC{$Γ ⟝ ⊤$}
\end{bprooftree}
\begin{bprooftree}
  \AxiomC{$Γ ⟝ φ ∨ ⊤$}
  \RightLabel{lem-ndisj$_{∨}$,}
  \UnaryInfC{$Γ ⟝ ⊤$}
\end{bprooftree}
\begin{bprooftree}
  \AxiomC{$Γ ⟝ φ ∨ ⊥$}
  \RightLabel{lem-ndisj$_{∨}$.}
  \UnaryInfC{$Γ ⟝ φ$}
\end{bprooftree}
\end{equation*}

\begin{equation*}
\begin{bprooftree}
  \AxiomC{$Γ ⟝ φ ∧ ¬~φ$}
  \RightLabel{lem-ndisj$_{∧}$,}
  \UnaryInfC{$Γ ⟝ ⊥$}
\end{bprooftree}
\begin{bprooftree}
  \AxiomC{$Γ ⟝ φ ∧ ⊤$}
  \RightLabel{lem-ndisj$_{∧}$,}
  \UnaryInfC{$Γ ⟝ φ$}
\end{bprooftree}
\begin{bprooftree}
  \AxiomC{$Γ ⟝ φ ∧ ⊥$}
  \RightLabel{lem-ndisj$_{∧}$.}
  \UnaryInfC{$Γ ⟝ ⊥$}
\end{bprooftree}
\end{equation*}

\begin{definition}[ndisj$_{∨}$]
  \label{def:ndisj-or}
  \begin{align*}
    \begin{split}
    &ndisj_{∨} : \NProp \to \NProp\\
    &ndisj_{∨}(φ) =
      \begin{cases}
        ⊤, &\text{if }φ ≡ ¬ φ₁ ∨ φ₂,\ φ₁ ∈_{∨} φ₂\\
        ⊤, &\text{if}φ ≡ ¬ φ₁ ∨ φ₂,\ ndisj_{∨}(φ₂) ≡ ⊤\\
        ¬ φ₁ ∨ ndisj_{∨}(φ₂), &\text{if }φ ≡ ¬ φ₁ ∨ φ₂\\
        ⊤, &\text{if }φ ≡ φ₁ ∨ φ₂,\ ¬φ₁ ∈_{∨} φ₂ \\
        ⊤, &\text{if }φ ≡ φ₁ ∨ φ₂,\ ndisj_{∨}(φ₂) ≡ ⊤\\
        φ₁ ∨ ndisj_{∨}(φ₂), &\text{if }φ ≡ φ₁ ∨ φ₂\\
        φ, &\text{otherwise.}
      \end{cases}
    \end{split}
  \end{align*}
\end{definition}

\begin{lemma}[lem-ndisj$_{∨}$] % (fold)
  \label{lem:lem_ndisj-or}
  $Γ ⟝ φ$ then $Γ ⟝ ndisj_{∨}(φ)$.
\end{lemma}

\begin{definition}[ndisj$_{∧}$]
  \label{def:ndisj-and}
  \begin{align*}
    \begin{split}
    &ndisj_{∧} : \NProp \to \NProp\\
    &ndisj_{∧}(φ) =
      \begin{cases}
        ⊥, &\text{ if }φ ≡ ¬ φ₁ ∧ φ₂,\ φ₁ ∈_{∧} φ₂\\
        ⊥, &\text{ if }φ ≡ ¬ φ₁ ∧ φ₂,\ ndisj_{∧}(φ₂) ≡ ⊥\\
        ¬ φ₁ ∧ ndisj_{∧}(φ₂), &\text{ if }φ ≡ ¬ φ₁ ∧ φ₂\\
        ⊥, &\text{ if }φ ≡ φ₁ ∧ φ₂,\ φ₁ ∈_{∧} φ₂\\
        ⊥, &\text{ if }φ ≡ φ₁ ∧ φ₂,\ ndisj_{∧}(φ₂) ≡ ⊥\\
        φ₁ ∧ ndisj_{∧}(φ₂), &\text{if }φ ≡ φ₁ ∧ φ₂\\
        φ, &\text{otherwise.}
      \end{cases}
    \end{split}
  \end{align*}
\end{definition}

\begin{lemma}[lem-ndisj$_{∧}$] % (fold)
  \label{lem:lem_ndisj-and}
  $Γ ⟝ φ$ then $Γ ⟝ ndisj_{∧}(φ)$.
\end{lemma}

% subsubsection redundancy (end)
% -------------------------------------------------------------------

We are ready to combine all function defined above in two function
$canon_{∧}$ for removing redundancies in conjunctions, and
$canon_{∨}$ for removing redundancies in disjunctions.

\begin{definition}[canon$_{∧}$]
\label{def:canon-and}
\begin{align*}
    \begin{split}
      &canon_{∧} : \NProp \to \NProp\\
      &canon_{∧}(φ) =
        \begin{cases}
        φⱼ, &\text{if } φ ≡ φ₁ ∧ φ₂,\ φᵢ ≡ ⊤,\text{ for } i,j = 1,2, i≠j\\
        ⊥, &\text{if } φ ≡ φ₁ ∧ φ₂,\ φᵢ ≡ ⊥,\text{ for } i,j = 1,2\\
       rm_∧(ndisj_∧(φ)), &\text{if } φ ≡ φ₁ ∧ φ₂\\
       φ, &\text{otherwise.}
        \end{cases}
    \end{split}
\end{align*}
\end{definition}

\begin{lemma}[lem-canon$_{∧}$]
  \label{lem:lem_canon-and}
  $Γ ⟝ φ$ then $Γ ⟝ canon_{∧}(φ)$.
\end{lemma}
